Ground state energy of quantum dots using the coupled cluster method

### 2018-04-17 project_2/article/article.tex
Include appendix on normal ordered Hamiltonian

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \sum_{pqrs}w_{rs}^{pq}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            w_{rs}^{pq} \equiv \bra{pq}v\ket{rs}.
        \end{gather}
        Note that the two-body matrix elements are not antisymmetric yet.

@@

    We start by writing the one-body operator, $h$, to its normal-ordered form.
    \begin{align}
        h &= \sum_{pq}h^p_q\acr{p}\ade{q}
        = \sum_{pq}h^p_q\para{
            \{\acr{p}\ade{q}\}
            + \{
                \wick{\c a_{p}^{\dagger} \c a_{q}}
            \}
        }
        \\
        &= \sum_{pq}h^p_q\{
            \acr{p}\ade{q}
        \}
        + \sum_{pq}h^p_q\delta_{p \in i}\delta_{pq}
        \\
        &= h_N + \sum_{i}h^i_i,
    \end{align}
    where we have used $\delta_{p \in i}$ to mean that $p$ must be an occupied
    index. Doing the same for the two-body operator is a slightly more tedious
    endeavor. For brevity we will only write out the operator strings and only
    keep the non-zero contributions.
    \begin{align}
        \acr{p}\acr{q}\ade{s}\ade{r}
    \end{align}
    \cite{crawford2007introduction}

>>>>

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            \bra{pq}\ket{rs} \equiv \bra{pq}w\ket{rs} - \bra{pq}w\ket{sr}.
        \end{gather}
        Note that we use the chemists notation to label the antisymmetric matrix
        elements.

@@

    We start by writing the one-body operator, $h$, to its normal-ordered form.
    \begin{align}
        h &= \sum_{pq}h^p_q\acr{p}\ade{q}
        = \sum_{pq}h^p_q\para{
            \{\acr{p}\ade{q}\}
            + \{
                \wick{\c a_{p}^{\dagger} \c a_{q}}
            \}
        }
        \\
        &= \sum_{pq}h^p_q\{
            \acr{p}\ade{q}
        \}
        + \sum_{pq}h^p_q\delta_{p \in i}\delta_{pq}
        \\
        &= h_N + \sum_{i}h^i_i,
    \end{align}
    where we have used $\delta_{p \in i}$ to mean that $p$ must be an occupied
    index. Doing the same for the two-body operator is a slightly more tedious
    endeavor. For brevity we will only write out the operator strings and only
    keep the non-zero contributions.
    \begin{align}
        \acr{p}\acr{q}\ade{s}\ade{r}
        &=
        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
        + \{
            \wick{\c a_p^{\dagger} a_{q}^{\dagger} \c a_{s} a_{r}}
        \}
        + \{
            \wick{\c a_p^{\dagger} a_{q}^{\dagger} a_{s} \c a_{r}}
        \}
        \nonumber \\
        &\qquad
        + \{
            \wick{a_p^{\dagger} \c a_{q}^{\dagger} \c a_{s} a_{r}}
        \}
        + \{
            \wick{a_p^{\dagger} \c a_{q}^{\dagger} a_{s} \c a_{r}}
        \}
        \nonumber \\
        &\qquad
        + \{
            \wick{\c1 a_p^{\dagger} \c2 a_{q}^{\dagger} \c1 a_{s} \c2 a_{r}}
        \}
        + \{
            \wick{\c2 a_p^{\dagger} \c1 a_{q}^{\dagger} \c1 a_{s} \c2 a_{r}}
        \}
        \\
        &=
        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
        - \delta_{p \in i}\delta_{ps}\{\acr{q}\ade{r}\}
        + \delta_{p \in i}\delta_{pr}\{\acr{q}\ade{s}\}
        \nonumber \\
        &\qquad
        + \delta_{q \in i}\delta_{qs}\{\acr{p}\ade{r}\}
        - \delta_{q \in i}\delta_{qr}\{\acr{p}\ade{s}\}
        \nonumber \\
        &\qquad
        - \delta_{p \in i}\delta_{ps}\delta_{q \in j}\delta_{qr}
        + \delta_{p \in i}\delta_{pr}\delta_{q \in j}\delta_{qs}.
    \end{align}
    Inserted into the full two-body operator and sorting out the sums we get
    \begin{align}
        W
        &=
        \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
        - \frac{1}{4}\sum_{iqr}\bra{iq}\ket{ri}\{\acr{q}\ade{r}\}
        \nonumber \\
        &\qquad
        + \frac{1}{4}\sum_{iqs}\bra{iq}\ket{is}\{\acr{q}\ade{s}\}
        + \frac{1}{4}\sum_{pir}\bra{pi}\ket{ri}\{\acr{p}\ade{r}\}
        \nonumber \\
        &\qquad
        - \frac{1}{4}\sum_{pis}\bra{pi}\ket{is}\{\acr{p}\ade{s}\}
        - \frac{1}{4}\sum_{ij}\bra{ij}\ket{ji}
        \nonumber \\
        &\qquad
        + \frac{1}{4}\sum_{ij}\bra{ij}\ket{ij}.
    \end{align}
    Using the antisymmetric properties of the two-body matrix elements,
    \begin{align}
        \bra{pq}\ket{rs}
        = - \bra{pq}\ket{sr}
        = - \bra{qp}\ket{rs}
        = \bra{qp}\ket{sr},
    \end{align}
    and relabeling of the indices we can rearrange and collect some terms.
    \begin{align}
        W
        &=
        W_N + \sum_{pir}\bra{pi}\ket{ri}\{\acr{p}\ade{r}\}
        + \half\sum_{ij}\bra{ij}\ket{ij},
    \end{align}
    where the normal ordered two-body operator is
    \begin{align}
        W_N = \frac{1}{4}\sum_{pqrs}
        \bra{pq}\ket{rs}\{\acr{p}\acr{q}\ade{s}\ade{r}\}.
    \end{align}
    When we now construct the full Hamiltonian we can collect some terms. The
    constants in both the one-body and the two-body operator in total constitues
    the reference energy.
    \begin{align}
        \bslat H\kslat
        = \sum_{i}h_i^i + \frac{1}{2}\sum_{ij}\bra{ij}\ket{ij}.
    \end{align}
    Combining the normal ordered one-body operator and the second term in the
    two-body operator, i.e., the term with a single creation and annihilation
    operator pair, we get the normal ordered Fock-operator.
    \begin{align}
        F_N
        &=
        \sum_{pq}h_{q}^{p}\{\acr{p}\ade{q}\}
        + \sum_{pqi}\bra{pi}\ket{qi}\{\acr{p}\ade{q}\}
        \\
        &= \sum_{pq}f_{q}^{p}\{\acr{p}\ade{q}\},
    \end{align}
    where we have defined the Fock matrix elements as
    \begin{align}
        f_{q}^{p}
        &=
        h_q^p + \sum_{i}\bra{pi}\ket{qi}.
    \end{align}
    In total we get the full Hamiltonian
    \begin{align}
        H
        &=
        F_N + W_N + \bslat H\kslat
        \\
        &= H_N + \bslat H\kslat,
    \end{align}
    which is what we wanted to show.\cite{crawford2007introduction}
